Ground state energy of quantum dots using the coupled cluster method

### 2018-06-07 project_2/article/article.tex

\section{Theory}
    In this project we will study a system of $N$ interacting electrons. We will
    be looking at a Hamiltonian consisting of a one-body and a two-body part.
    The one-body part is given by
    \begin{align}
        h(\vf{r}_i)
        &= -\half\nabla_i^2 + \half\omega^2 \vf{r}_i^2,
    \end{align}
    where we use natural units $\hbar = c = e = 1$ and set the mass to unity.
    % Why not write $\hbar = c = e = m = 1$?
    The two-body part is the Coulomb interaction potential.
    \begin{align}
        u(\vf{r}_i, \vf{r}_j)
        &= \frac{1}{\abs{\vf{r}_i - \vf{r}_j}}.
    \end{align}
    We thus get the total Hamiltonian
    \begin{align}
        H &= h + u
        =
        \sum_{i = 1}^N h(\vf{r}_i) + \sum_{i < j}^N u(\vf{r}_i, \vf{r}_j),
    \end{align}
    where $h$ is the full one-body operator and $u$ the full two-body
    operator, i.e., over the entire system.  Working in a basis of $L$ single
    particle functions, $\{\ket{p}\}_{p = 1}^L$. We define the reference Slater
    determinant as
    \begin{align}
        \ket{\Phi_0} &\equiv \ket{1, 2, \dots, N},
    \end{align}
    i.e., a tensorproduct of the $N$ first single particle functions, $\ket{i}$,
    of the system. We call these single particle functions \emph{occupied} as
    they are contained in the Slater determinant.  We will denote the occupied
    indices with $i, j, k, l, \dots \in \{1, \dots, N\}$, the \emph{virtual}
    states with $a, b, c, d, \dots \in \{N + 1, \dots, L\}$ and general indices
    with $p, q, r, s, \dots \in \{1, \dots, L\}$. In terms of sets of basis
    functions we can write this as
    \begin{align}
        \{\ket{p}\}_{p = 1}^L
        = \{\ket{i}\}_{i = 1}^N
        \cup
        \{\ket{a}\}_{a = N + 1}^L,
    \end{align}
    i.e., the general indexed states consists of both occupied and virtual
    states. Note that the single particle functions are orthonormal, i.e.,
    \begin{align}
        \braket{p}{q} = \delta_{pq}.
    \end{align}
    We can construct other Slater determinants in this basis by exciting
    or relaxing the reference determinant. A general excitation is labeled
    $\ket{\Phi_{ij\dots}^{ab\dots}}$ which means that we have removed the single
    particle functions with indices $i, j, \dots$ from the reference and added
    $a, b, \dots$. Note that
    \begin{align}
        \braket{\Phi^{ab\dots}_{ij\dots}}{\Phi_0} = 0,
        \label{eq:excited_overlap}
    \end{align}
    for any excitation.

>>>>

\section{Theory}
    In this project we will study a system of $N$ interacting electrons. We will
    be looking at a Hamiltonian consisting of a one-body and a two-body part.
    The one-body part is given by
    \begin{align}
        h(\vf{r}_i)
        &= -\half\nabla_i^2 + \half\omega^2 \vf{r}_i^2,
        \label{eq:one_body_hamiltonian}
    \end{align}
    where we use natural units $\hbar = c = e = 1$ and set the mass to unity.
    % Why not write $\hbar = c = e = m = 1$?
    The two-body part is the Coulomb interaction potential,
    \begin{align}
        u(\vf{r}_i, \vf{r}_j)
        &= \frac{1}{\abs{\vf{r}_i - \vf{r}_j}}.
        \label{eq:two_body_operator}
    \end{align}
    We thus get the total Hamiltonian
    \begin{align}
        H &= h + u
        =
        \sum_{i = 1}^N h(\vf{r}_i) + \sum_{i < j}^N u(\vf{r}_i, \vf{r}_j),
    \end{align}
    where $h$ is the full one-body operator and $u$ the full two-body
    operator, i.e., over the entire system.  Working in a basis of $L$ single
    particle functions, $\{\ket{p}\}_{p = 1}^L$. We define the reference Slater
    determinant as
    \begin{align}
        \ket{\Phi_0} &\equiv \ket{1, 2, \dots, N},
    \end{align}
    i.e., a tensorproduct of the $N$ first single particle functions, $\ket{i}$,
    of the system. We call these single particle functions \emph{occupied} as
    they are contained in the Slater determinant.  We will denote the occupied
    indices with $i, j, k, l, \dots \in \{1, \dots, N\}$, the \emph{virtual}
    states with $a, b, c, d, \dots \in \{N + 1, \dots, L\}$ and general indices
    with $p, q, r, s, \dots \in \{1, \dots, L\}$. In terms of sets of basis
    functions we can write this as
    \begin{align}
        \{\ket{p}\}_{p = 1}^L
        = \{\ket{i}\}_{i = 1}^N
        \cup
        \{\ket{a}\}_{a = N + 1}^L,
    \end{align}
    i.e., the general indexed states consists of both occupied and virtual
    states. Note that the single particle functions are orthonormal, i.e.,
    \begin{align}
        \braket{p}{q} = \delta_{pq}.
    \end{align}
    We can construct other Slater determinants in this basis by exciting
    or relaxing the reference determinant. A general excitation is labeled
    $\ket{\Phi_{ij\dots}^{ab\dots}}$ which means that we have removed the single
    particle functions with indices $i, j, \dots$ from the reference and added
    $a, b, \dots$. Note that
    \begin{align}
        \braket{\Phi^{ab\dots}_{ij\dots}}{\Phi_0} = 0,
        \label{eq:excited_overlap}
    \end{align}
    for any excitation.

    \subsection{Quantum Dots}

        Some consideration should be given towards why the system described above
        is a quantum dot, and why we should care about quantum dots.
        There are at least two things we refer to as quantum dots. The
        first one is a man-made device, also known as artificial atoms,
        and the other is the simplest theoretical construct of such a device.

        Aritifical atom quantum dots are semiconducting crystals, typically GaAs
        relatively easy to build in a laboratory. Here, some free electrons are 
        confined to a volume, either by a physical barrier like an insulator
        or an electromagnetic field\cite{reimann2002electronic}. Due to the
        confinement of the electrons, the energy levels of a quantum dot becomes
        quantized, and they therefore hold many properties similar to naturally
        occuring quantized systems, which is the reason why quantum dots have
        attracted much interest.

        The theoretical approximation of a quantum dot is impossible to make perfectly
        accurate, yet simple. There are some important effects that should be
        included such the potential arising as a result of the confinement, and
        the interaction between particles. We therefore model our system with an
        electron-electron interaction, given by the two-body operator in
        \autoref{eq:eq:two_body_operator}, and with we approximate the confining
        potential by the harmonic oscillator potential, given by the second term
        in the one-body Hamiltonian in \autoref{eq:eq:one_body_hamiltonian}. The
        last term in the one-body Hamiltonian is the kinetic energy of particle
        in question, essential in any system.
